Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)

### 2024-02-21 manuscript/revision2/revision2.qmd
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on both genetic variation and environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions. Identifying the mixture and prevalence of GxE primed by weather prior to the phenological event (GxWeather) should aid prediction and manipulation of phenological timings. Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of these populations at eight sites spanning the central United States. We distinguish GxWeather patterns covarying with interpretable weather-based cues from agnostic, site-based patterns. Most GxWeather effects belong to the latter category and do not covary with weather-based cues. However, 65% of effects on the timing of vegetative growth in the Gulf population covary with daylength 14 days prior to green-up date. 33% of effects on the timing of flowering in the Gulf population covary with cumulative rainfall in the seven days prior to flowering, while 22% of effects on flowering timing in the Midwest population covary with day length change two days prior to flowering. An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped 23 additive QTLs for flowering, all with significant associations and eleven with enrichment of highly significant associations in our model of GxWeather effects. We demonstrate that we can identify QTL with GxWeather and assign them to specific weather-based cues or other patterns. Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod and rainfall cues in switchgrass. More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.

@@

  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep fitness cost of plant-environment mismatches. We infer the mixture of ways genetic effects on phenological traits depend on plant environment, focusing on how effects covary with weather prior to the phenological event (GxWeather). Most effects do not have GxWeather, but a minority do. GxWeather is population-specific. The majority of effects on the timing of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue two weeks prior. A minority of effects on flowering date in the Gulf and Midwest populations had GxWeather, covarying with a cumulative rainfall and a photoperiod cue, respectively.

@@

\firstpage[9]{2}

Plant phenological timings are major components of plant fitness affected by multiple external environmental cues (e.g. degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology]. Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed. Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year. However, the usefulness of specific environmental cues depends on both features of the environment, such as cue predictability and relevance, and the species' adaptive strategies @botero2015evolutionary. Species with wide natural distributions can have multiple distinct environmentally cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions]. Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.

@@

Switchgrass (*Panicum virgatum*) is considered a short-day plant with reproductive development strongly linked to day of the year @mitchell1997predicting. However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal]. We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and survival @lovell2021genomic. The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the well-studied lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021genomic.

Here, we estimate the ways genetic effects on phenological timings depend on weather prior to the phenological event. To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude. These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters. We define multiple ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance types that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible. We use the Bayesian framework *mash* (multivariate adaptive shrinkage), developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites. *mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest. Importantly, *mash* does not have a statistical bias in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041. To confirm our genetic mapping of GxWeather, we compare our posterior genetic effect estimates from *mash* to mapping results from an outbred pseudo-F2 cross grown at the same sites. Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.

@@

![Caption for this figure... blah blah blah.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}

![Comparisons of pairs of jointly re-estimated effects across eight gardens, for effects with lfsr \< 0.05. AHM is considering moving this figure to the supplement and substantially simplifying the figure in the main text, perhaps just showing total effects in Texas, the North, and maybe Oklahoma. Essentially making Table 2 into a new, simplified figure. ](images/Figure_3_Effect_Types_by_Site.png){#fig-effects}

>>>>

  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on both genetic variation and environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions. Identifying the mixture and prevalence of GxE primed by weather prior to the phenological event (GxWeather) should aid prediction and manipulation of phenological timings. Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites spanning the central United States. We distinguish GxWeather patterns covarying with interpretable weather-based cues from agnostic, site-based patterns. Most GxWeather effects belong to the latter category and do not covary with weather-based cues. However, 65% of effects on the timing of vegetative growth in the Gulf population covary with daylength 14 days prior to green-up date. 33% of effects on the timing of flowering in the Gulf population covary with cumulative rainfall in the seven days prior to flowering, while 22% of effects on flowering timing in the Midwest population covary with day length change two days prior to flowering. An independent pseudo-F2 cross of Gulf and Midwest individuals at the same sites mapped 10 additive QTLs for flowering and three for vegetative growth timing, all with significant associations and three with enrichment of highly significant associations in our model of GxWeather effects. We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns. Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod and rainfall cues in switchgrass. More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.

@@

  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep fitness cost of plant-environment mismatches. We infer the mixture of ways that genetic effects on phenological traits depend on plant environment, focusing on how effects covary with weather prior to the phenological event (GxWeather). Most effects do not have GxWeather, but a minority do. GxWeather is population-specific. The majority of effects on the timing of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue two weeks prior. A minority of effects on flowering date in the Gulf and Midwest populations had GxWeather, covarying with a cumulative rainfall and a photoperiod cue, respectively.

@@

\firstpage[8]{2}

Plant phenological timings are major components of plant fitness affected by multiple external environmental cues (e.g. degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions [@bauerle_photoperiodic_2012; @andres2012genetic; @korner2010phenology]. Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed. Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year. However, the usefulness of specific environmental cues depends on both features of the environment, such as cue predictability and relevance, and the species' adaptive strategies @botero2015evolutionary. Species with wide natural distributions can have multiple distinct environmentally cued phenological responses: for example, populations of sunflower (*Helianthus annuus*) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments [@blackman2013interacting; @henry2014transitions]. Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.

@@

Previous research suggests that switchgrass phenological timings should have GxWeather and that these timings could differ by genetic subpopulation. Switchgrass is considered a short-day plant with reproductive development strongly linked to day of the year @mitchell1997predicting. However, as part of its wide environmental adaptation across the eastern half of North America, its photoperiodicity has been predicted to differ by plant latitude of origin [@parrish2005biology; @casler2004latitudinal]. We previously found divergent Midwest and Gulf genetic subpopulations of switchgrass with distinct sets of environmental adaptations, in that both populations had distinct genetic variation associated with each of two fitness proxies, biomass and survival @lovell2021genomic. The Midwest genetic subpopulation is primarily composed of individuals from the well-studied upland switchgrass ecotype [@porter1966analysis; @milano2016genetic], while the Gulf subpopulation has individuals from the well-studied lowland ecotype and the phenotypically intermediate coastal ecotype @lovell2021genomic.

Here, we estimate GxWeather for two phenological timings in switchgrass by assigning patterns of genetic effects on phenology across gardens to many patterns of weather covariance at these gardens. To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude. These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters. We define multiple ways phenological timings might covary with weather (\@tbl-covar) and additional ways phenological timings might vary by site (\@SI-info), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (\@SI-info) @urbut2019flexible. We use the Bayesian framework *mash* (multivariate adaptive shrinkage), developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites. *mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest. Importantly, *mash* does not have a statistical bias in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041. To confirm our genetic mapping of GxWeather, we compare our posterior genetic effect estimates from *mash* to mapping results from an outbred pseudo-F2 cross grown at the same sites. Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.

@@

![Example hypothesis-driven covariance matrices specified in mash and the posterior weights placed on all covariance matrices. (a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table XXX. Common gardens are arranged in latitudinal order within the matrices. Right column: Five example canonical covariance matrices. Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden. (b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date mash models, within and between two genetic subpopulations. Hypothesized covariance matrices (green). Covariance matrices included in mash that had zero posterior weight in all three mash runs on the genetic subpopulations, such as the identity matrix, are not shown. (c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_Covariance_matrices_and_posterior_weights.png){#fig-covar}

![Comparisons of pairs of jointly re-estimated effects across eight gardens, for effects with lfsr \< 0.05. AHM is considering moving this figure to the supplement and substantially simplifying the figure in the main text, perhaps just showing total effects in Texas, the North, and maybe Oklahoma. Essentially making Table 2 into a new, simplified figure. Effect patterns for re-estimated SNP effects that are significant at pairs of sites by the local false sign rate test (lfsr), for (a) greenup date and (b) flowering date, within and between two genetic subpopulations. Common gardens are arranged in latitudinal order along the x- and y-axis. Effects in the first and fourth rows differ in sign at these pairs of gardens (p \< 0.05, lfsr). Effects in the second, third, fifth and sixth rows are identical in sign (p \< 0.05, lfsr). Effects in the second and fifth rows additionally differ in magnitude by a factor of \>0.4, while effects in the third and sixth row were not distinguishable by magnitude nor sign of the effect.](images/Figure_3_Effect_Types_by_Site.png){#fig-effects}

### 2024-02-21 manuscript/revision2/revision2.tex
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

\significancestatement{The timing of plant seasonal development
(phenology) has major impacts on fitness because of the steep fitness
cost of plant-environment mismatches. We infer the mixture of ways
genetic effects on phenological traits depend on plant environment,
focusing on how effects covary with weather prior to the phenological
event (GxWeather). Most effects do not have GxWeather, but a minority
do. GxWeather is population-specific. The majority of effects on the
timing of vegetative growth in the Gulf population have GxWeather: these
effects differ in sign across environments and covary with a photoperiod
cue two weeks prior. A minority of effects on flowering date in the Gulf
and Midwest populations had GxWeather, covarying with a cumulative
rainfall and a photoperiod cue, respectively.}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on both genetic variation and
environmental cues; thus, phenological traits can have important
genotype-by-environment (GxE) interactions. Identifying the mixture and
prevalence of GxE primed by weather prior to the phenological event
(GxWeather) should aid prediction and manipulation of phenological
timings. Here, we map GxWeather effects on phenological timings in two
highly divergent switchgrass (\emph{Panicum virgatum}) populations using
repeated plantings of these populations at eight sites spanning the
central United States. We distinguish GxWeather patterns covarying with
interpretable weather-based cues from agnostic, site-based patterns.
Most GxWeather effects belong to the latter category and do not covary
with weather-based cues. However, 65\% of effects on the timing of
vegetative growth in the Gulf population covary with daylength 14 days
prior to green-up date. 33\% of effects on the timing of flowering in
the Gulf population covary with cumulative rainfall in the seven days
prior to flowering, while 22\% of effects on flowering timing in the
Midwest population covary with day length change two days prior to
flowering. An independent pseudo-F2 cross of Gulf and Midwest
individuals at the same sites mapped 23 additive QTLs for flowering, all
with significant associations and eleven with enrichment of highly
significant associations in our model of GxWeather effects. We
demonstrate that we can identify QTL with GxWeather and assign them to
specific weather-based cues or other patterns. Breeding for particular
alleles at these loci could change flowering responsiveness to
photoperiod and rainfall cues in switchgrass. More broadly, this
approach could be used to identify genetic marker-environment
interactions in any species with related populations phenotyped in
multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[9]{2}

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (Helianthus annuus) exhibit day-neutral, facultative short
day, and facultative long-day flowering responses, which vary with their
environments (5, 6). Distinct genetic responses in different
environments are known as genotype by environment interactions, or GxE.

@@

Switchgrass (\emph{Panicum virgatum}) is considered a short-day plant
with reproductive development strongly linked to day of the year (29).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (30, 31). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and survival (32). The Midwest genetic subpopulation is primarily
composed of individuals from the well-studied upland switchgrass ecotype
(33, 34), while the Gulf subpopulation has individuals from the
well-studied lowland ecotype and the phenotypically intermediate coastal
ecotype (32).

Here, we estimate the ways genetic effects on phenological timings
depend on weather prior to the phenological event. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(@tbl-covar) and additional ways phenological timings might vary by site
(@SI-info), then jointly re-estimate genetic effects on these timings at
all eight sites using the set of these covariance types that
significantly improved the modeled log-likelihood when included
(@SI-info) (35). We use the Bayesian framework \emph{mash} (multivariate
adaptive shrinkage), developed by (35), to refine effect size estimates
from genome-wide association (GWAS) conducted on individuals at each of
the eight sites. \emph{mash} allows us to identify and specify multiple
covariance structures among genetic effect estimates across sites,
including structures that represent covariance in weather variables of
interest. Importantly, \emph{mash} does not have a statistical bias in
detecting genetic effects with the same or opposite signs (36). To
confirm our genetic mapping of GxWeather, we compare our posterior
genetic effect estimates from \emph{mash} to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
to describe the weather cues and genetic variation affecting phenology
in two divergent natural populations of switchgrass.

@@

\caption{\label{fig-covar}Caption for this figure\ldots{} blah blah
blah.}

@@

\caption{\label{fig-effects}Comparisons of pairs of jointly re-estimated
effects across eight gardens, for effects with lfsr \textless{} 0.05.
AHM is considering moving this figure to the supplement and
substantially simplifying the figure in the main text, perhaps just
showing total effects in Texas, the North, and maybe Oklahoma.
Essentially making Table 2 into a new, simplified figure.}

>>>>

\significancestatement{The timing of plant seasonal development
(phenology) has major impacts on fitness because of the steep fitness
cost of plant-environment mismatches. We infer the mixture of ways that
genetic effects on phenological traits depend on plant environment,
focusing on how effects covary with weather prior to the phenological
event (GxWeather). Most effects do not have GxWeather, but a minority
do. GxWeather is population-specific. The majority of effects on the
timing of vegetative growth in the Gulf population have GxWeather: these
effects differ in sign across environments and covary with a photoperiod
cue two weeks prior. A minority of effects on flowering date in the Gulf
and Midwest populations had GxWeather, covarying with a cumulative
rainfall and a photoperiod cue, respectively.}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on both genetic variation and
environmental cues; thus, phenological traits can have important
genotype-by-environment (GxE) interactions. Identifying the mixture and
prevalence of GxE primed by weather prior to the phenological event
(GxWeather) should aid prediction and manipulation of phenological
timings. Here, we map GxWeather effects on phenological timings in two
highly divergent switchgrass (\emph{Panicum virgatum}) populations using
repeated plantings of cloned individuals from these populations at eight
sites spanning the central United States. We distinguish GxWeather
patterns covarying with interpretable weather-based cues from agnostic,
site-based patterns. Most GxWeather effects belong to the latter
category and do not covary with weather-based cues. However, 65\% of
effects on the timing of vegetative growth in the Gulf population covary
with daylength 14 days prior to green-up date. 33\% of effects on the
timing of flowering in the Gulf population covary with cumulative
rainfall in the seven days prior to flowering, while 22\% of effects on
flowering timing in the Midwest population covary with day length change
two days prior to flowering. An independent pseudo-F2 cross of Gulf and
Midwest individuals at the same sites mapped 10 additive QTLs for
flowering and three for vegetative growth timing, all with significant
associations and three with enrichment of highly significant
associations in our model of GxWeather effects. We demonstrate that we
can identify genetic variation with GxWeather and assign these loci to
specific weather-based cues or other patterns. Breeding for particular
alleles at these loci could change flowering responsiveness to
photoperiod and rainfall cues in switchgrass. More broadly, this
approach could be used to identify genetic marker-environment
interactions in any species with related populations phenotyped in
multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[8]{2}

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
short day, and facultative long-day flowering responses, which vary with
their environments (5, 6). Distinct genetic responses in different
environments are known as genotype by environment interactions, or GxE.

@@

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (29).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (30, 31). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and survival (32). The Midwest genetic subpopulation is primarily
composed of individuals from the well-studied upland switchgrass ecotype
(33, 34), while the Gulf subpopulation has individuals from the
well-studied lowland ecotype and the phenotypically intermediate coastal
ecotype (32).

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology across gardens to
many patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(@tbl-covar) and additional ways phenological timings might vary by site
(@SI-info), then jointly re-estimate genetic effects on these timings at
all eight sites using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included
(@SI-info) (35). We use the Bayesian framework \emph{mash} (multivariate
adaptive shrinkage), developed by (35), to refine effect size estimates
from genome-wide association (GWAS) conducted on individuals at each of
the eight sites. \emph{mash} allows us to identify and specify multiple
covariance structures among genetic effect estimates across sites,
including structures that represent covariance in weather variables of
interest. Importantly, \emph{mash} does not have a statistical bias in
detecting genetic effects with the same or opposite signs (36). To
confirm our genetic mapping of GxWeather, we compare our posterior
genetic effect estimates from \emph{mash} to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
to describe the weather cues and genetic variation affecting phenology
in two divergent natural populations of switchgrass.

@@

\caption{\label{fig-covar}Example hypothesis-driven covariance matrices
specified in mash and the posterior weights placed on all covariance
matrices. (a) Left column: Five example hypothesized covariance matrices
specified for the green-up date or flowering date phenotype; these
matrices were created from environment-specific correlations across
eight common gardens, and are described in Table XXX. Common gardens are
arranged in latitudinal order within the matrices. Right column: Five
example canonical covariance matrices. Canonical matrices (purple) have
simple interpretations, such as equal effects across all common gardens,
or effects specific to a single common garden. (b,d) Total posterior
weight placed on each covariance matrix type specified for (b) green-up
date and (d) flowering date mash models, within and between two genetic
subpopulations. Hypothesized covariance matrices (green). Covariance
matrices included in mash that had zero posterior weight in all three
mash runs on the genetic subpopulations, such as the identity matrix,
are not shown. (c,e) Total posterior weight placed on covariance
matrices that were hypothesized or canonical, for the (c) green-up date
phenotype and (e) flowering date phenotype.}

@@

\caption{\label{fig-effects}Comparisons of pairs of jointly re-estimated
effects across eight gardens, for effects with lfsr \textless{} 0.05.
AHM is considering moving this figure to the supplement and
substantially simplifying the figure in the main text, perhaps just
showing total effects in Texas, the North, and maybe Oklahoma.
Essentially making Table 2 into a new, simplified figure. Effect
patterns for re-estimated SNP effects that are significant at pairs of
sites by the local false sign rate test (lfsr), for (a) greenup date and
(b) flowering date, within and between two genetic subpopulations.
Common gardens are arranged in latitudinal order along the x- and
y-axis. Effects in the first and fourth rows differ in sign at these
pairs of gardens (p \textless{} 0.05, lfsr). Effects in the second,
third, fifth and sixth rows are identical in sign (p \textless{} 0.05,
lfsr). Effects in the second and fifth rows additionally differ in
magnitude by a factor of \textgreater0.4, while effects in the third and
sixth row were not distinguishable by magnitude nor sign of the effect.}
